---
title: "COMP2501 Assignment 1"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Requirements

**Submission deadline: Feb 27th, 2025 at 23:59.**

**Full mark of assignment 1: 50.**

For the following questions, please:

1. Replace all [Input here] places with your information or your answer (for multiple choice).
2. Complete the code block by adding your own code to fulfill the requirements in each question. Please use the existing code block and do not add your own code block.

Please make sure your Rmd file is a valid Markdown document and can be successfully knitted.

For assignment submission, please knit your final Rmd file into a Word document, and submit both your **Rmd** file and the knitted **Microsoft Word** document file to Moodle. You get 0 score if 1) the Rmd file you submitted cannot be knitted, and 2) you have not submitted a Word document. For each visualization question, please make sure that the generated plot is shown in-place with the question and after the code block. 

---


## Name and UID

Name: [Tianhong Gao]

UID: [3036459300]

---


### Environmental setup 
You need to have the `dslabs`, `dplyr`, and `ggplot2` packages installed. If not yet, please run `install.packages(c("dslabs", "dplyr", "ggplot2"))` in your R environment. If you have installed the `tidyverse` package, `dplyr` is installed by default.

```{r}

# Load the required packages
# install.packages(c("dslabs", "dplyr", "ggplot2"))
library(dslabs)
library(dplyr)
library(ggplot2)

```


### Example question. Print the first 2 records of the murders dataset.

```{r}

data("murders")
head(murders, 2)

```


### 1. (1 points) Given matrix `mat`, print the entries of row 2, and columns 2 to 4. Please define a matrix of your choice. You will need to select both row 2 and columns 2 to 4 simultaneously.

```{r}

mat<-matrix(1:20,nrow=4,ncol=5)
mat[2, 2:4]


```


### 2. (2 points) Write a function `compute_s_n`  that for any given n, computes the $S_n = n*\sqrt{(n+9)}*log_{10}(n)$. Print the $S_n$ with $n = 500$.

```{r}

compute_s_n <- function(n){
  n*sqrt(n+9)*log10(n)
}
compute_s_n(500)


```   


### 3. (1 points)  Examine the built-in dataset `Orange`. Which of the following is true?
- a.	Orange is tidy data: it has one observation for each row.
- b.	Orange is not tidy: we need at least one column with a character vector.
- c.	Orange is not tidy: it is a matrix instead of a data frame.
- d.	Orange is tidy data: all small datasets are tidy by definition.

Your answer is: [a]


### 4. (12 points) Utilize the murders dataset and execute the following tasks.

```{r}

# Load the required packages and dataset
library(dslabs)
library(dplyr)
library(ggplot2)
data("murders")

```

#### a. (2 points) Compute the murder rate per 100,000 people for each state and store it in an object called murder_rate. Then use logical operators to find which state has a murder rate per 100,000 people higher than 5. Find these states, print their names and murder rate per 100,000 people.

```{r}

murders<-mutate(murders,rate=total/population*100000)
filt_rate<-filter(murders,rate>5)
result<-select(filt_rate, state, rate)
print(result)

```


#### b. (2 points) For all states having a murder rate per 100,000 people higher than 5, use the barplot function to create a barplot with the x-axis being the state name, and the y-axis being the murder rate per 100,000 people of each state. (Hints: check some barplot examples at https://r-graph-gallery.com/210-custom-barplot-layout.html)

```{r}

barplot(filt_rate$rate, names.arg=filt_rate$state, xlab="State", ylab="Murder rate per 100000 people higher than 5")

```


#### c. (3 points) Create a table called `my_states` that contains rows for states satisfying two conditions: 1) it is in either West or South, and 2) the murder rate per 100,000 people is less than 2.0. Use `select` to show only the state name, the region and the murder rate, and use `top_n` function to find the 3 safest states among them.

```{r}


```


#### d. (2 points) Compute the average murder rate per 100,000 people in the four regions respectively of the U.S., and sort the results by murder rate in ascending order.

```{r}


```


#### e. (3 points) Use the ggplot2 package to create a scatterplot from the murders dataset, where the x-axis is the number of population, the y-axis is the total number of murders, and each point in the scatterplot is labeled with the state name. Please add an appropriate title, and axis labels to the plot.

```{r}


```


### 5. (4 points)  Utilize the mtcars dataset and `ggplot2` package to execute the following tasks.

```{r}

# Load the required packages and dataset.
library(ggplot2)
data("mtcars")

```

#### a. (2 points) Generate a scatter plot with ggplot2 showing the relationship between car weight (wt) and miles per gallon (mpg). Include a title and axis labels in your plot.

```{r}


```


#### b. (2 points) Create a box plot that compares the distribution of miles per gallon (mpg) across different numbers of gears (gear). Customize the colors of the boxes and add a title and axis labels.

```{r}


```


### 6. (7 points) Analyze the diamonds dataset to identify the impact of different factors on diamond prices.

```{r}

# Load the required packages and dataset
library(dplyr)
library(ggplot2)
data("diamonds")

```

#### a. (3 points) Use scatterplot to display how cut, color, and clarity affect the price of a diamond. Ensure that the axes are appropriately labeled and that each type of cut, color, and clarity is distinguishable via colored labels. (Hints: You could display how cut, color, and clarity affect the price of a diamond separately, which means you could draw three separate scatterplots. And you could think about considering carat as the independent variable for the x-axis, and cut, color, and clarity as the colored labels separately to draw the three scatterplots successfully.)

```{r}


```


#### b. (2 points) Group the diamonds dataset based on their clarity attribute. For every group, clearly present the average depth of the diamonds as well as average cost per carat weight of that particular clarity of diamond. Additionally, for every group, show the name of the most frequently occurring color code (D - J). Ensure that the results are sorted based on the frequency of the clarity group observed.

```{r}


```


#### c. (2 points) Make use of an appropriate graph (a density plot or a line chart) to show the distribution of diamond prices in the diamonds dataset.

```{r}


```


### 7. (23 points)  Explore the tidyverse with the COVID-19 dataset (http://www.bio8.cs.hku.hk/comp2501/covid.csv), and answer the following questions. Noticeably, it is acceptable to analyze the data or generate a plot with negative values of cases and deaths.

```{r}

# Load the required packages
library(dplyr)
library(ggplot2)

```

#### a. (2 points) Read the CSV formatted dataset. Find out how many observations (rows) and variables (columns) are in the dataset. Print the names of all variables.

```{r}


```


#### b. (1 points) List the observation with the largest `Cumulative_number_for_14_days_of_COVID.19_cases_per_100000`.

```{r}


```


#### c. (2 points) How many unique `countriesAndTerritories` are in the dataset? How many unique `continentExp` are in the dataset?

```{r}


```


#### d. (3 points) For 1) the whole dataset, 2) different `countriesAndTerritories`, and 3) different `continentExp`, compute both i) the sum of cases, and ii) the sum of deaths. Sort the results by the sum of cases descendingly. Use head() if there are too many rows in the results. 

```{r}


```


#### e. (2 points) Add a new column `date` with the standard date format "YYYY-MM-DD" to the data table according to the `dateRep` column. Be reminded the format of `dateRep` is "DD/MM/YYYY". Please use head() to show the result.

```{r}


```


#### f. (1 points) Create a scatterplot showing `cases` vs. `deaths`. Set an appropriate plot title and axis titles.

```{r}


```


#### g. (2 points) Create a line plot using data with `countriesAndTerritories=="China"`, showing date on the x-axis and cases per day on the y-axis. Set an appropriate plot title and axis titles.

```{r}


```


#### h. (2 points) Similar to above, create a line plot using the data of six countries including "China", "United_States_of_America", "United_Kingdom", "France", "Germany", and "Italy". Use different line colors for each country. Set an appropriate plot title and axis titles.

```{r}


```


#### i. (4 points) Filter out the data in Germany. Assume the Germany government determined an epidemic situation of national significance on 2020-06-10. A lot of restrictive measurements such as social distancing were implemented to prevent the spread of COVID-19. Find out the 1) average number of cases before the determination, 2) average number of cases after the determination, 3) average number of deaths before the determination, and 4) average number of deaths after the determination. Remove missing values from data if they are coded as NA, if there is any.

```{r}


```


#### j. (4 points) In fact, the determination of epidemic situation was on another date. Make a reasonable guess with explanation by the following steps. Create a line plot using data with `countriesAndTerritories=="Germany"`, showing date on the x-axis, cases per day and death per day on the y-axis. Observe the pattern and make a reasonable guess. Find a way to indicate the date of the determination and 1 month after the date of determination. Set an appropriate plot title and axis titles. Marks would be given on answers with reasonable explanations. (Hints: It is expected that the cases confirmed would drop after 1 month of the implementation of determination of epidemic situation, with 14-week incubation period of COVID-19. It is a relatively open question, and reasonable explanations are acceptable. We wanna you to observe the trend of changes in the cases per day and the deaths per day on the date you have chosen and one month after it according to the line plot you created. No correct date is needed and we will judge your answers according to your thoughts and explanations.)

```{r}


```

